Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-21 introduction.tex
lloid_paper: one more typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1125 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updated figure 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1127 ccdd05d5-56c1-461a-bf98-85f6876523b4

\begin{table}[h]
\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for sources that produce an \SNR\ of 8 in all three detectors at time $t$ before merger.  The fifth and sixth columns show the minimum area of the 90\% confidence region before merger and at merger.}
\begin{center}
\begin{tabular}{rrrrrr}
\tableline\tableline
& horiz. & rate & final & \multicolumn{2}{c}{$A$(90\%) (deg$^2$)} \\
\cline{5-6}
$t$ (s) & (Mpc) & (yr$^{-1}$) & \SNR\ & early & final \\
\tableline
25.0 & 226 & 5 & 15.8 & 2893.2 & 1.8 \\
10.0 & 299 & 12 & 11.9 & 1150.6 & 3.6 \\
5.0 & 341 & 18 & 10.4 & 594.6 & 5.1 \\
1.0 & 401 & 29 & 8.9 & 153.9 & 7.5 \\
0.1 & 435 & 37 & 8.2 & 32.5 & 9.1 \\
0.0 & 445 & 40 & 8.0 & 9.6 & 9.6 \\
\tableline
\end{tabular}
\end{center}
\end{table}
%
\begin{figure}[h]
\plotone{figures/loc_in_time.pdf}
\caption{\label{fig:sky-localization-accuracy} Area of the 90\% confidence
region as a function of time for an exceptional source, where such a source is
defined to produce a signal with an \SNR\ of at least 27.4. These signals are
expected to be observed in the detectors once per year.  80~s, 35~s, 16~s, and
6~s before coalescence, such a signal will have accumulated an \SNR\ of 8, 12,
16, and 20, respectively. After merger, the minimum 90\% confidence area
shrinks to $0.82$~deg$^2$, shown by the dashed line.}
\end{figure}

@@

It is infeasible to search this hundreds of square degrees for a prompt
counterpart. For
reference, \textsc{lsst} should see first light around the same time as
Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
\citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
to accurately reflect mass and thus be a proxy for the \CBC{} event rate, it is
possible to reduce the localization uncertainty using a galaxy catalog, along
with the sky location and luminosity distance estimate from the \GW{} signal,
as a prior probability distribution \citep{galaxy-catalog} as was done in
S6/VSR3. Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
distance, the number of galaxies that can produce a given signal
amplitude is much larger than the number for initial \LIGO{} and thus the
catalog will not be as useful for downselecting pointings generally. However,
our exceptional sources will necessarily be extremely nearby. Within this
reduced volume there will be fewer galaxies to consider for a given
candidate and catalog completeness will be less of a concern.
This should allow us to shrink the 90\% confidence area substantially.

>>>>

\begin{table}[h]
\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for sources that produce an \SNR\ of 8 in all three detectors at time $t$ before merger.  The fifth and sixth columns show the minimum area of the 90\% confidence region before merger and at merger.}
\begin{center}
\begin{tabular}{rrrrrr}
\tableline\tableline
& horiz. & rate & final & \multicolumn{2}{c}{$A$(90\%) (deg$^2$)} \\
\cline{5-6}
$t$ (s) & (Mpc) & (yr$^{-1}$) & \SNR\ & early & final \\
\tableline
25.0 & 226 & 5 & 15.8 & 2893.2 & 1.8 \\
10.0 & 299 & 12 & 11.9 & 1150.6 & 3.6 \\
5.0 & 341 & 18 & 10.4 & 594.6 & 5.1 \\
1.0 & 401 & 29 & 8.9 & 153.9 & 7.5 \\
0.1 & 435 & 37 & 8.2 & 32.5 & 9.1 \\
0.0 & 445 & 40 & 8.0 & 9.6 & 9.6 \\
\tableline
\end{tabular}
\end{center}
\end{table}
%
\begin{figure}[h]
\plotone{figures/loc_in_time.pdf}
\caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
region as a function of time for a sources with anticipated detection rates
of 10~yr$^{-1}$, 1~yr$^{-1}$, and 0.1~yr$^{-1}$. The heavy dot indicates the
time at which the accumulated \SNR\ exceeds a threshold of 8.}
\end{figure}

@@

It is infeasible to search hundreds of square degrees for a prompt
counterpart. For
reference, \textsc{lsst} should see first light around the same time as
Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
\citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
to accurately reflect mass and thus be a proxy for the \CBC{} event rate, it is
possible to reduce the localization uncertainty using a galaxy catalog, along
with the sky location and luminosity distance estimate from the \GW{} signal,
as a prior probability distribution \citep{galaxy-catalog} as was done in
S6/VSR3. Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
distance, the number of galaxies that can produce a given signal
amplitude is much larger than the number for initial \LIGO{} and thus the
catalog will not be as useful for downselecting pointings generally. However,
our exceptional sources will necessarily be extremely nearby. Within this
reduced volume there will be fewer galaxies to consider for a given
candidate and catalog completeness will be less of a concern.
This should allow us to shrink the 90\% confidence area substantially.
